---
output: github_document
---

```{r echo=FALSE, results = 'asis'}
pkg <- "rRDP"

source("https://raw.githubusercontent.com/mhahsler/pkg_helpers/main/pkg_helpers.R")
pkg_title(pkg, CRAN = FALSE, Bioc = FALSE)
```

**Note:** RDP has aged and it was time 
end support and to remove the package from Bioconductor. 

The package is still functional and can be installed via:

```{r eval = FALSE}
install.packages(c('rRDP', 'rRDPData'), repos = c('https://mhahsler.r-universe.dev'))
```


This package installs and interfaces the naive Bayesian classifier for 16S rRNA 
sequences developed by the Ribosomal Database Project (RDP). With this package the 
classifier trained with the standard training set can be used or a custom classifier can be trained.

## Usage

```{r, message=FALSE}
library(rRDP)
```

Load a few sequences and classify the sequences using RDP.

```{r}
seq <- readRNAStringSet(system.file("examples/RNA_example.fasta",
  package="rRDP"))
seq

pred <- predict(rdp(), seq)
pred

attr(pred, "confidence")
```


## Citation Request

```{r, echo=FALSE, results='asis', comment= ""}
pkg_citation(pkg)
```

## Acknowledgments
This work was partially supported by grant no. R21HG005912 from the [National Human Genome Research Institute](https://www.genome.gov/).

