---
output: github_document
---

```{r echo=FALSE, results = 'asis'}
pkg <- "rRDP"

source("https://raw.githubusercontent.com/mhahsler/pkg_helpers/main/pkg_helpers.R")
pkg_title(pkg, CRAN = FALSE, Bioc = FALSE)
```




This package installs and interfaces the naive Bayesian classifier for 16S rRNA 
sequences developed by the Ribosomal Database Project (RDP). With this package the 
classifier trained with the standard training set can be used or a custom classifier can be trained.

## Installation

```{r eval = FALSE}
if (!require("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}

BiocManager::install("Biostrings")

install.packages(c('rRDP'), repos = c('https://mhahsler.r-universe.dev'))
```


## Usage

```{r, message=FALSE}
library(rRDP)
```

Load a few sequences and classify them with RDP. On first use, `rdp()` asks
before downloading the default 16S classifier to the user data directory.
The download may take a while.

```{r}
seq <- readRNAStringSet(system.file("examples/RNA_example.fasta",
  package="rRDP"))
seq

pred <- predict(rdp(), seq)
pred

attr(pred, "confidence")
```


## Citation Request

This package was formerly distributed via Bioconductor and the updated version here 
should still be cited as: 

```{r, echo=FALSE, results='asis', comment= ""}
pkg_citation(pkg)
```

## Acknowledgments
This work was partially supported by grant no. R21HG005912 from the [National Human Genome Research Institute](https://www.genome.gov/).
