---
title: "rRDP: Interface to the RDP Classifier"
author: "Michael Hahsler and Anurag Nagar"
bibliography: rRDP.bib
vignette: >
  %\VignetteIndexEntry{rRDP: Interface to the RDP Classifier}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
abstract: "This package installs and interfaces the naive Bayesian classifier for 16S rRNA sequences developed by the Ribosomal Database Project (RDP). With this package the classifier trained with the standard training set can be used or a custom classifier can be trained."
output:
  rmarkdown::html_vignette
---


```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

```{r setup}
library("rRDP")
set.seed(1234)
```

# Installation and System Requirements

rRDP requires the Bioconductor package Biostrings and R to be configured with Java.

```{r, eval=FALSE}
if (!require("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}

BiocManager::install("Biostrings")
```

Install rRDP and the database used by the default RDP classifier.

```{r, eval=FALSE}
install.packages(c('rRDP'), repos = c('https://mhahsler.r-universe.dev'))
```

RDP uses Java and you need a working installation of the `rJava` package.
You need to have a Java JDK installed. On Linux, you can install Open JDK
and run in your shell 
`R CMD javareconf` to configure R for using Jave.
On Windows, you can install the latest version of the JDK from 
https://www.oracle.com/java/technologies/downloads/ and set the 
`JAVA_HOME` environment variable in R using (make sure to use the correct location).
An example would look like this:

```{r, eval=FALSE}
Sys.setenv(JAVA_HOME = "C:\\Program Files\\Java\\jdk-20")
```

Note the double backslashes (i.e. escaped slashes) used in the path.

Details 
can be found at https://www.rforge.net/rJava/index.html.
To configure R for Java, 

## How to cite this package

```{r, comment = ""}
citation("rRDP")
```

# Classification with RDP
The RDP classifier was developed by the Ribosomal Database Project [@rdp:Cole:2003]
which provides various tools and services to the scientific community for data 
related to 16S rRNA sequences. The classifier uses a Naive Bayesian approach to 
quickly and accurately classify sequences. The classifier uses 
8-mer counts as features [@rdp:Wang:2007].

## Training a custom RDP classifier

RDP can be trained using `trainRDP()`. We use an example of a tiny training 
data set that is shipped with the package.

```{r}
trainingSequences <- readDNAStringSet(
    system.file("examples/trainingSequences.fasta", package = "rRDP")
)
trainingSequences
```

Note that the training data needs to have names in a specific RDP format: 

`"<ID> <Kingdom>;<Phylum>;<Class>;<Order>;<Family>;<Genus>"`


In the following we show the name for the first sequence.
We use here `sprintf` to display
only the first 65~characters so the it fits into a single line.

```{r}
sprintf(names(trainingSequences[1]), fmt = "%.65s...")
```


Now, we can train a the classifier. The model is stored in a directory specified
by the parameter `dir`.

```{r}
customRDP <- trainRDP(trainingSequences, dir = "myRDP")
customRDP
```

As test sequences, we sample 5 sequences form the training data and extract a 
subsequences of length 500. 
```{r}
testSequences <- sample(trainingSequences, 5)

testSequences <- subseq(testSequences, start = 500, width = 500)
testSequences
```

```{r}
pred <- predict(customRDP, testSequences)
pred
```

The prediction confidence is supplied as the attribute `"confidence"`.
```{r}
attr(pred, "confidence")
```

To evaluate the classification accuracy we can compare the known classification
with the predictions. The known classification was stored in the FASTA file
and encoded in RDP format. We can decode the annotation using 
`decode_RDP()`. Greengenes format can be decoded using `decode_Greengenes()`.

```{r}
annotation <- names(testSequences)
actual <- decode_RDP(annotation)
actual
```
  
Now we can compare the prediction with the actual classification by creating
a confusion table and calculating the classification accuracy. Here we do this
at the Genus level.

```{r}
confusionTable(actual, pred, rank = "genus")
accuracy(actual, pred, rank = "genus")
```

Note that the confidence and accuracy values are not very informative since we trained the 
classifier with a tiny data set so the vignette can be created fast. You can use 
a larger training set or a pretrained classifier to get more informative results.

  
Since the custom classifier is stored on disc it can be recalled anytime using
`rdp()`.

```{r}
customRDP <- rdp(dir = "myRDP")
customRDP
```


To permanently remove the classifier use `removeRDP()`. This will delete the 
directory containing the classifier files.

```{r}
removeRDP(customRDP)
```

\subsection{Using the RDP classifier trained with the default training set}
RDP is trained with a 16S rRNA training set. 
The default classifier is downloaded on first use and cached in the user's
R data directory. It uses RDP Classifier 2.14 released in August 2023,
with the bacterial and archaeal taxonomy training set No. 19
[@rdp:Wang:2024].

Since the download take a while, it is not shown executed in this vignette.

```{r, eval=FALSE}
default_16S_classifier <- rdp()
```


# Session Info
```{r}
sessionInfo()
```

# Acknowledgments
This research is supported by research grant no. R21HG005912 from the National Human Genome Research Institute (NHGRI / NIH).

# References
